# Reproducible Research - Peer Assessment 1

*Libraries used*
```{r,message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
```

**Loading and preprocessing the data**

Data loaded (imho no further preprocessing is needed)
```{r}
data <- read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric"))
head(data)
```

**What is mean total number of steps taken per day?**

Total number of steps taken per day calculated (only part of the data is shown)
```{r}
steps_per_day <- group_by(data, date) %>% summarize(steps = sum(steps, na.rm=TRUE))
head(steps_per_day)
```

Histogram made
```{r}
qplot(steps_per_day$steps, main = "Histogram of the total number of steps taken each day", xlab = "steps per day")
```

Mean and median calculated
```{r}
mean <- mean(steps_per_day$steps)
median <- median(steps_per_day$steps)
mean
median
```
The mean of the total number of steps taken per day is `r mean` , the median is `r median`.

**What is the average daily activity pattern?**

Plot made
```{r}
average_steps <- group_by(data, interval) %>% summarize(avrg_steps = mean(steps, na.rm=TRUE))
qplot(average_steps$interval, average_steps$avrg_steps, geom = "line", 
      main = "Average number of steps taken per time interval", 
      xlab = "time interval", ylab = "average number of steps")
```

Interval with maximum number of steps calculated
```{r}
max <- as.numeric(average_steps[average_steps$avrg_steps == max(average_steps$avrg_steps),1])
max
```
The interval with maximum number of steps (averaged aross all the days) is `r max`.

**Imputing missing values**

Total number of missing values calculated
```{r}
NAs <- sum(is.na(data$steps))
NAs
```
Total number of missing values in the dataset is `r NAs`.

New dataset with imputed missing values created. Mean for the relevant 5-minute interval was used.
```{r}
data_with_avrg <- arrange(merge(data, average_steps, by = 'interval'), date, interval)
data_with_avrg$steps[is.na(data_with_avrg$steps)]=data_with_avrg$avrg_steps[is.na(data_with_avrg$steps)]
noNA_data <- data_with_avrg[,1:3]
head(noNA_data)
```

Histogram made and mean and median calculated
```{r}
noNA_steps_per_day <- group_by(noNA_data, date) %>% summarize(steps = sum(steps, na.rm=TRUE)) 
qplot(noNA_steps_per_day$steps, main = "Histogram of the total number of steps taken each day", xlab = "steps per day")
noNA_mean <- mean(noNA_steps_per_day$steps)
noNA_median <- median(noNA_steps_per_day$steps)
noNA_mean
noNA_median
```
After the imputation of missing values, the mean of the total number of steps taken per day is `r noNA_mean` , the median is `r noNA_median`. They differ from the values calculated for dataset with missing values, they are slightly higher.

**Are there differences in activity patterns between weekdays and weekends?**

New factor variable created
```{r}
noNA_data <- mutate(noNA_data, wd_we=weekdays(noNA_data$date, abbreviate = TRUE))
noNA_data$wd_we <- gsub("(po)|(út)|(st)|(čt)|(pá)", "weekday", noNA_data$wd_we)
noNA_data$wd_we <- gsub("(so)|(ne)", "weekend", noNA_data$wd_we)
noNA_data$wd_we <- as.factor(noNA_data$wd_we)
```

Panel plot made
```{r}
steps_per_wd_we <- group_by(noNA_data, wd_we, interval) %>% summarize(steps = mean(steps)) 
qplot(interval, steps, data = steps_per_wd_we, geom = "line", facets = wd_we~.)
```
